\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}\label{ch:acknowledgements}
\vspace{-1cm}
The old saying \emph{where there's a will, there's a way} fails to mention that
proper support and means are must-haves in the journey that one takes for a PhD.
Many institutions and people have
a direct hand in the successful completion of this thesis.

I thank the EMBL predoc program, the University of Cambridge
and Hughes Hall college to have provided me
with the working and living environment
for one of the most formative periods of my life.

I am grateful to my supervisor \alvis\
without whom this adventure wouldn't have happened.
Thank you, Alvis, for accepting me into the Functional Genomic group
which comprises many exceptional people
that contributed to my day-to-day life and enriched me as a person.
Thank you for your trust and support
through this roller coaster that was my doctorate.

I am thankful to \jyoti\ and \james\ for their collaboration.
I wouldn't have ever delved into the proteomics world
if it wasn't for their guidance and discussions.
I also would like to thank my thesis advisory committee,
Dr Sarah Teichman, Dr Gos Micklem and Dr Wolfgang Huber,
for their expert feedback and discussion.
Many thanks to Lynn French, Lorraine McAlister, Anna Alasalmi
and Clare Impey for helping with many administrative tasks
and paperwork involved in my doctorate.\mybr\

I can't thank my teammates enough for all the welcoming,
exciting and fun environment.
Thank you for all your help and the lunch conversations about anything
from top-notch science to absurd elephant jokes.
Thanks to Mar, Nuno, Johan, Wanseon, Sérgio, Claudia, Aida (Fatemeh)
for their discussions and close friendship.
Special thanks to Nuno and Claudia,
who provided me with the support I needed at crucial times.
Nuno, thank you for your invaluable work and help with processing the data.
You are like a bigger brother with whom I could discuss anything,
from the inner hell circle of R to SciFi, in the most in-depth tiny details.
Obrigada Nuno for all the years we share at the EBI.\@

I am most grateful to the EBI community and
particularly to the EBI predocs who are amazing people.
Thanks to our morning coffee discussions and debates,
I learnt so much about bioinformatic fields that
were not directly involved in my work.
I am truly fortunate to have met all of you.
Furthermore, I want to thank more specifically the fantastic people
who have taken time to read and give me feedback on this thesis.
Mar, Nuno, Juan, Myrto, Aida, Sarah, Julian, Konrad, Nils,
Hannah, Atoussa, Isabelle and Steve,
I thank all of you for your invaluable inputs and corrections.
Steve, thank you also so much for all the discussions
about the fine points in British culture and language.

Thank you Isabelle for our hour-long conversations,
which, for the most part, consisted of me figuring out the best ways
to explain my doctorate works to a layperson.
Thank you, Philippe, for your unfailing support and invaluable assistance,
particularly regarding the logistics.
Finally, I want to thank my sister, Atoussa, and my mother, Faranak,
to always be by my side, despite any possible physical distance.
Thank you both for helping me to grow and develop,
if not formalise, my ideas.
Thank you for sharing with me your own scientific and life interests,
which, beyond being exciting, sustain my joy and creativity.


